Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-09-10 Report/kkt_mse_approach.tex
finished convergence analysis

Even though the rate constraints are implicitly present in the objective function, we cannot formulate the \ac{KKT} conditions readily due to the non-differentiable objective function. The non-differentiability of the objective function is due to the absolute operator present in the norm function. In order to make the objective function differentiable, we consider the following case for which the absolute operator can be ignored without affecting the optimal solution, namely,
\begin{itemize}
\item when the exponent \me{q} is even or,
\item when the number of backlogged packets of each user is large enough \me{Q_k \gg \sum_{n=1}^N \sum_{l=1}^L {t}_{l,k,n}} to ignore the absolute operator and the queues as well.
\end{itemize}

With the assumption of either one of the above conditions to be true, the problem in \eqref{eqn-mse-2} can be written as
\begin{IEEEeqnarray}{rCl}\label{kkt-mse-1}
\underset{\substack{t_{l,k,n},\mvec{M}{k,n}, \\ \epsilon_{l,k,n}, \mvec{W}{k,n}}}{\text{minimize}} &\hspace{0.5cm}& \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \Big ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \Big )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
\text{subject to} & \hspace{0.5cm} & \nonumber \\
\alpha_{l,k,n} : & \hspace{0.5cm} & \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 +  N_0 \, \|\mvec{w}{l,k,n}\|^2 \nonumber \\
&\hspace{0.5cm}& \qquad {} \qquad {} + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,y,n} \mvec{m}{x,y,n} \right |^2 \leq \epsilon_{l,k,n} \IEEEyessubnumber \label{kkt-mse-1.2} \\
\sigma_{l,k,n} : & \hspace{0.5cm} & - \log(\tilde{\epsilon}_{l,k,n}) - \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\tilde{\epsilon}_{l,k,n}} \geq t_{l,k,n} \, \log(2) \IEEEyessubnumber \label{kkt-mse-1.3} \\
\delta_b : & \hspace{0.5cm} & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b \IEEEyessubnumber \label{kkt-mse-1.4}
\end{IEEEeqnarray}
where \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b} are the dual variables corresponding to the constraints defined in \eqref{kkt-mse-1.2}, \eqref{kkt-mse-1.3} and \eqref{kkt-mse-1.4}.

By forming the Lagrangian of \eqref{kkt-mse-1} with the corresponding dual variables as shown in \eqref{kkt-mse-1}, we can obtain the \ac{KKT} expressions by differentiating with respect to the variables present in the problem as detailed in Appendix \ref{a-1}. Upon solving the \ac{KKT} expressions, the iterative solution \me{\forall k \in \mc{U}}, \me{\forall n \in \set{1,\ldots,N}} and \me{\forall l \in \set{1,\ldots,L}} is given by
\begin{IEEEeqnarray}{rCl} \label{kkt-mse-4}
\mvec{m}{l,k,n}^{(i)} &=& \Big ( \sum_{x \in \mc{U}} \sum_{y=1}^L \alpha_{y,x,n}^{(i-1)} \mvec{H}{b_k,x,n}^\herm \mvec{w}{y,x,n}^{(i-1)} \mvec{w}{y,x,n}^{\herm \, {(i-1)}} \mvec{H}{b_k,x,n} + \delta_b \mbf{I}_{N_T} \Big )^{-1} \alpha^{(i-1)}_{l,k,n} \mvec{H}{b_k,k,n}^\herm \mvec{w}{l,k,n}^{(i-1)} \IEEEyessubnumber \label{kkt-mse-4.3} \\
\epsilon_{l,k,n}^{(i)} &=& \left | 1 - \mvec{w}{l,k,n}^{\herm \, (i)} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)} \right |^2 + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^{\mathrm{H} \, (i)} \mvec{H}{b_y,y,n} \mvec{m}{x,y,n}^{(i)} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}^{(i)}\|^2 \IEEEyessubnumber \label{kkt-mse-4.4} \\
t_{l,k,n}^{(i)} &=&  -\log_2(\epsilon_{l,k,n}^{(i-1)}) - \frac{\left ( \epsilon_{l,k,n}^{(i)} - \epsilon_{l,k,n}^{(i-1)} \right ) }{\log(2) \, \epsilon_{l,k,n}^{(i-1)}} \IEEEyessubnumber \label{kkt-mse-4.5} \\
\sigma_{l,k,n}^{(i)} &=& \max{\left (0, \frac{a_k \, q}{\log(2)}  \, \Big (Q_k - \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n}^{(i)} \Big )^{(q-1)}\right )}  \IEEEyessubnumber \label{kkt-mse-4.2} \\
\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \dfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1} \\
\mvec{w}{l,k,n}^{(i)} &=& \Big ( \sum_{x\in\mc{U}}\sum_{y=1}^L \mvec{H}{b_x,k,n} \mvec{m}{y,x,n}^{(i)} \mvec{m}{y,x,n}^{\herm \, (i)} \mvec{H}{b_{x},k,n}^\herm + \mathbf{I}_{N_R} \Big ) ^{-1} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}^{(i)} \IEEEyessubnumber \label{kkt-mse-4.6}
\end{IEEEeqnarray}
The dual variable \me{\alpha^{(i)}} is updated with memory as in \eqref{kkt-mse-4.1} to avoid abrupt oscillations due to \me{\sigma < 0} from \eqref{kkt-mse-4.2} when \me{Q_k < \sum_{n=1}^N \sum_{l=1}^L {t}_{l,k,n}}. The parameter \me{\rho \in [0,1]} provides a linear combination of the previous to the current value of the dual variable, dictating system reaction for the excess allocation.

>>>>

Even though the rate constraints are implicitly present in the objective function, we cannot formulate the \ac{KKT} conditions readily due to the non-differentiable objective function. The non-differentiability of the objective function is due to the absolute operator present in the norm function. In order to make the objective function differentiable, we consider the following case for which the absolute operator can be ignored without affecting the optimal solution, namely,
\begin{itemize}
\item when the exponent \me{q} is even or,
\item when the number of backlogged packets of each user is large enough, \textit{i.e}, \me{Q_k \gg \sum_{n=1}^N \sum_{l=1}^L {t}_{l,k,n}} to ignore the absolute operator and the queues as well.
\end{itemize}

With the assumption of either one of the above conditions to be true, the problem in \eqref{eqn-mse-2} can be written as
\begin{IEEEeqnarray}{rCl}\label{kkt-mse-1}
\underset{\substack{t_{l,k,n},\mvec{M}{k,n}, \\ \epsilon_{l,k,n}, \mvec{W}{k,n}}}{\text{minimize}} &\hspace{0.5cm}& \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \Big ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \Big )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
\text{subject to} & \hspace{0.5cm} & \nonumber \\
\alpha_{l,k,n} : & \hspace{0.5cm} & \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 +  N_0 \, \|\mvec{w}{l,k,n}\|^2 \nonumber \\
&\hspace{0.5cm}& \qquad {} \qquad {} + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,y,n} \mvec{m}{x,y,n} \right |^2 \leq \epsilon_{l,k,n} \IEEEyessubnumber \label{kkt-mse-1.2} \\
\sigma_{l,k,n} : & \hspace{0.5cm} & - \log(\tilde{\epsilon}_{l,k,n}) - \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\tilde{\epsilon}_{l,k,n}} \geq t_{l,k,n} \, \log(2) \IEEEyessubnumber \label{kkt-mse-1.3} \\
\delta_b : & \hspace{0.5cm} & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l = 1}^L \text{tr} \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}, \fall b \IEEEyessubnumber \label{kkt-mse-1.4}
\end{IEEEeqnarray}
where \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b} are the dual variables corresponding to the constraints defined in \eqref{kkt-mse-1.2}, \eqref{kkt-mse-1.3} and \eqref{kkt-mse-1.4}.

By forming the Lagrangian of \eqref{kkt-mse-1} with the corresponding dual variables as shown in \eqref{kkt-mse-1}, we can obtain the \ac{KKT} expressions by differentiating with respect to the variables present in the problem as detailed in Appendix \ref{a-1}. Upon solving the \ac{KKT} expressions, the iterative solution \me{\forall k \in \mc{U}}, \me{\forall n \in \set{1,\ldots,N}} and \me{\forall l \in \set{1,\ldots,L}} is given by
\begin{IEEEeqnarray}{rCl} \label{kkt-mse-4}
\mvec{m}{l,k,n}^{(i)} &=& \Big ( \sum_{x \in \mc{U}} \sum_{y=1}^L \alpha_{y,x,n}^{(i-1)} \mvec{H}{b_k,x,n}^\herm \mvec{w}{y,x,n}^{(i-1)} \mvec{w}{y,x,n}^{\herm \, {(i-1)}} \mvec{H}{b_k,x,n} + \delta_b \mbf{I}_{N_T} \Big )^{-1} \alpha^{(i-1)}_{l,k,n} \mvec{H}{b_k,k,n}^\herm \mvec{w}{l,k,n}^{(i-1)} \IEEEyessubnumber \label{kkt-mse-4.3} \\
\epsilon_{l,k,n}^{(i)} &=& \left | 1 - \mvec{w}{l,k,n}^{\herm \, (i)} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)} \right |^2 + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^{\mathrm{H} \, (i)} \mvec{H}{b_y,y,n} \mvec{m}{x,y,n}^{(i)} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}^{(i)}\|^2 \IEEEyessubnumber \label{kkt-mse-4.4} \\
t_{l,k,n}^{(i)} &=&  -\log_2(\epsilon_{l,k,n}^{(i-1)}) - \frac{\left ( \epsilon_{l,k,n}^{(i)} - \epsilon_{l,k,n}^{(i-1)} \right ) }{\log(2) \, \epsilon_{l,k,n}^{(i-1)}} \IEEEyessubnumber \label{kkt-mse-4.5} \\
\sigma_{l,k,n}^{(i)} &=& \max{\left (0, \frac{a_k \, q}{\log(2)}  \, \Big (Q_k - \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n}^{(i)} \Big )^{(q-1)}\right )}  \IEEEyessubnumber \label{kkt-mse-4.2} \\
\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \dfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1} \\
\mvec{w}{l,k,n}^{(i)} &=& \Big ( \sum_{x\in\mc{U}}\sum_{y=1}^L \mvec{H}{b_x,k,n} \mvec{m}{y,x,n}^{(i)} \mvec{m}{y,x,n}^{\herm \, (i)} \mvec{H}{b_{x},k,n}^\herm + \mathbf{I}_{N_R} \Big ) ^{-1} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}^{(i)} \IEEEyessubnumber \label{kkt-mse-4.6}
\end{IEEEeqnarray}
Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in the expressions in \eqref{kkt-mse-4.1}, one has to be fixed to optimize for the other. In this problem, we fix the dual variable \me{\alpha^{(i)}} to a fixed value as in \eqref{kkt-mse-4.1} to obtain the other variables in \eqref{kkt-mse-4}. It can be seen that, when the allocated rate \me{t_k^{(i-1)}} is greater than the number of queued packets \me{Q_k} for a user \me{k}, the corresponding dual variable \me{\sigma^{(i)}} will be negative thereby forcing the dual variable \me{\alpha_k^{(i)} < \alpha_k^{(i-1)}}, which in turn reduce the transmit precoder weights in \eqref{kkt-mse-4.3}. This reduction in the precoder weights forces the allocated rate of the user \me{k} to reduce from the previous iteration as \me{t_k^{(i)} < t_k^{(i-1)}}.

### 2014-09-10 Report/mse_formulation.tex
finished convergence analysis

\subsubsection*{Convergence}
The convergence of the \ac{MSE} reformulation approach follows the same arguments detailed in \cite{wmmse_shi,hong2012decomposition}. Since the transmit precoders are obtained by minimizing the weighted \ac{MSE} criterion, it converges to the stationary \ac{KKT} point at each iteration. The receiver updates are made based on the \ac{MMSE} criterion, it provides the optimal receive beamformers for the given transmit precoders. Alternating between these two precoder designs guarantees the monotonicity of the objective function used in the problem \eqref{eqn-mse-2}. The \ac{MSE} relaxation method converges to the same stationary \ac{KKT} point as of the original \ac{WSRM} problem by making the gradients equal.

>>>>

\subsubsection*{Convergence}
Following the similar argument from Section \ref{sec-3.2.1}, we can see that the \ac{SCA} subproblems converges to a stationary point. The transmit precoders obtained by the \ac{MSE} relaxations are unique due to the \ac{MSE} expression in \eqref{eqn-mse-2.3}, since it removes the phase ambiguity due to the real valued expression for the transmit precoders as
\begin{equation}
\epsilon_{l,k,n} \geq  1 - 2 \Re{ \left \lbrace \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right \rbrace} + \sum_{\mathclap{\forall (j,i)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2
\end{equation}
where \me{\Im{ \left \lbrace \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right \rbrace}} is zero due to the real valued \ac{MSE}. Since the transmit precoders are obtained by minimizing the weighted \ac{MSE} criterion, it converges to the stationary \ac{KKT} point at each iteration. The receiver updates are made based on the \ac{MMSE} criterion, it provides the optimal receive beamformers for the given transmit precoders. Alternating between these two precoder designs guarantees the monotonicity of the objective function used in the problem \eqref{eqn-mse-2}. The \ac{MSE} relaxation method converges to the same stationary \ac{KKT} point as of the original \ac{WSRM} problem by making the gradients equal.

### 2014-09-10 Report/wsrm_formulation.tex
finished convergence analysis

In the proposed solution, we replaced \eqref{eqn-6.2} by a convex constraint using the first order approximation, which basically means that we do not solve the problem exactly. According to the traditional \ac{BCDM}, we need to solve a sub problem when fixing a set of variables to the global optimum to ensure the convergence to a stationary point. If we just approximate the objective, then the convergence is guaranteed \cite{razaviyayn2013unified}. In our case, we solve the sub problem inexactly, so the convergence proof of \ac{BCDM} does not apply to our problem. In this problem, we used alternating optimization with \ac{SCA} method, which provides monotonic convergence since the objective is improved at each step \textit{i.e} \me{f^{(i)} \geq f^{(i+1)}}, assuming \me{f^{(i)}} is the objective function at the \me{i^\mathrm{th}} \ac{SCA} iteration. For a fixed receive beamformers, the problem is guaranteed to converge to a \ac{KKT} point at each \ac{SCA} updates as discussed in \cite{marks1978technical}. Now, by alternating the transmit and the receive precoders, we perform \ac{BCDM}, which provides monotonic convergence of the objective.

The linear approximation majorizes the quadratic-over-linear function in \eqref{eqn-6.2} around a fixed point \me{\tilde{\mbf{u}}^{(i)}_{l,k,n}}. The constraint approximation satisfies the following conditions as in \cite{marks1978technical}
\begin{subeqnarray} \label{sca-req}
	f(\tilde{\mbf{u}}_{l,k,n}) &\leq& \bar{f}(\tilde{\mbf{u}}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	\partial f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \partial \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n})
\end{subeqnarray}
where \me{\bar{f}(x,x^{(i)})} is the approximate function majorizing \me{\bar{f}(x)} around the point \me{x^{(i)}}. At each iteration \me{(i)}, the optimal value \me{x^{\ast}} obtained by solving the approximate problem is a feasible solution to next approximation at \me{(i+1)} iteration by following \eqref{sca-req}. Once the problem converged to an optimal point, the receive beamformers are updated based on the receivers mentioned in \eqref{opt-rx} or \eqref{eqn-10}. Since the receiver minimizes the objective value for the fixed transmit precoders, the proposed \ac{JSFRA} scheme guarantees to converge to an optimal point.

>>>>

In order to prove the convergence of the proposed algorithm, we require the following conditions to be satisfied
\begin{itemize}
	\item convergence of the \ac{SCA} subproblem	
	\item uniqueness of the transmit and the receive beamformers obtained
	\item monotonic convergence of the objective function
\end{itemize}
In the proposed solution, we replaced \eqref{eqn-6.2} by a convex constraint using the first order approximation, the linear approximation is majorized by the quadratic-over-linear function in \eqref{eqn-6.2} from below around a fixed point \me{\tilde{\mbf{u}}^{(i)}_{l,k,n}}. Since the \ac{SCA} method is adopted in the proposed algorithm, the constraint approximation satisfies the following conditions as in \cite{marks1978technical}
\begin{subeqnarray} \label{sca-req}
	f(\tilde{\mbf{u}}_{l,k,n}) &\leq& \bar{f}(\tilde{\mbf{u}}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n}) \\
	\partial f(\tilde{\mbf{u}}^{(i)}_{l,k,n}) &=& \partial \bar{f}(\tilde{\mbf{u}}^{(i)}_{l,k,n},\tilde{\mbf{u}}^{(i)}_{l,k,n})
\end{subeqnarray}
where \me{\bar{f}(x,x^{(i)})} is the approximate function \me{\bar{f}(x)} around the point \me{x^{(i-1)}}. At each iteration \me{(i-1)}, the optimal value \me{x^{\ast}} obtained by solving the approximate problem is a feasible solution to next approximation at \me{(i)} iteration by following \eqref{sca-req}. It can be easily verified that the optimal point of each \ac{SCA} problem satisfies the \ac{KKT} condition of the original problem, which in fact is the stationary point of the \ac{SCA} subproblem.

The uniqueness of the transmit and the receive beamformers can be justified by forcing one antenna to be real valued to exclude the phase ambiguity arising from the complex precoders. The monotonic convergence of the objective function can be justified by the following arguments. At each \ac{SCA} iteration, the relaxed subproblem is solved for the optimal transmit precoders to minimize the objective function. Since the \ac{SCA} subproblem is relaxed around the \me{\ith{(i-1)}} optimal point for the \me{\ith{(i)}} iteration, the solution of the \me{\ith{(i)}} iteration includes optimal point from the  \me{\ith{(i-1)}} iteration as well. Therefore, at each \ac{SCA} iteration, the objective function can either be equal to or smaller than the previous solution, thereby leading to the monotonic convergence of the objective function.

Once the problem converged to an optimal point, the receive beamformers are updated based on the receivers mentioned in \eqref{opt-rx} or \eqref{eqn-10}. Since the receiver minimizes the objective value for the fixed transmit precoders, the proposed \ac{JSFRA} scheme guarantees to converge to an optimal point.
